Prediction of Early Periprosthetic Joint Infection after Total Hip Arthroplasty: Development and External Validation of a Multivariable Model Based on Observational Registry data from Sweden and Denmark

### 2022-01-15 reports/article.Rmd
initial submission to clinep

---
title: "Prediction of Early Periprosthetic Joint Infection after Total Hip Arthroplasty: Development and External Validation of a Multivariable Model Based on Observational Registry data from Sweden and Denmark"
author: Erik Bülow,^1,2^ Ute Hahn,^3,4^ Ina Trolle Andersen,^3^ Ola Rolfson,^1,2^  Alma Besic Pedersen,^3,5^ Nils P. Hailer^6^
date: '`r Sys.Date()`'
bibliography: 
#  H:/zotero_lib.bib
  C:\\Users\\erik_\\Documents\\library.bib
csl: bmj.csl
output: 
  bookdown::word_document2:
    reference_docx: article.dotm
    toc: false
    df_print: kable
    number_sections: false
---

@@

- EB: Statistician, MSc, PhD, erik.bulow@gu.se
- ITA: Statistician, MSc, PhD, ita@clin.au.dk
- UH: Biostatistician, PhD, associate professor, ute@math.au.dk
- OR: Orthopaedic consultant, PhD, professor, register director, ola.rolfson@gu.se
- ABP: Orthopaedic consultant, PhD, DMSc, professor, abp@clin.au.dk
- NPH: Orthopaedic consultant, PhD, professor, nils.Hailer@surgsci.uu.se

@@

**Word count:** XXX

@@

# Summary box

## What is already known on this topic

- Periprosthetic joint infection (PJI) is the most devastating of early complications after total hip arthroplasty (THA).
- Attempts have been made to develop individualized risk prediction models for clinical usage in the US and Australia but none of those models has gained general acceptance due to limited samples, low discriminating ability, sub-optimal calibration or absence of external model validation.

## What this study adds
- Our prediction model was derived on a Swedish cohort and externally validated on a Danish cohort
- Our model is based on easily accessible data that are available before THA
- Compared to more complex comorbidity indices or to ASA class, our model had superior discriminatory ability and good calibration
- We provide a simple web calculator for easy use in clinical practice (https://erikbulow.shinyapps.io/thamortpred/).

@@

## Patient and Public Involvement

Patient representatives are elected as ordinary members of the steering committees for SHAR and DHR, where the development of shared decision making has been discussed. Our web calculator will be available for (potential) patients and the public.

@@

## Conclusion

@@

@@

# Disclosure

@@

```{r comparecoefs}
compare_se_dk_coefs %>% 
  rename(`95 % CI` = CI) %>% 
knitr::kable(
  caption = "Estimated coefficients for the reduced model based on the Swedish derivation cohort (Swedish), as well as re-estimated coefficient values based on the Danish cohort (Danish). Empirical confidence intervals (95 % CI) based on 1,000 resamples from the Swedish cohort of the same sample size (N = 18,854) as used in the Danish cohort. The Danish estimates fall within most of the CIs, indicating no support to reject the null hypothesis of no differneces between the countries."
)
```

>>>>

---
title: "Prediction of Early Periprosthetic Joint Infection after Total Hip Arthroplasty: Development and External Validation of a Multivariable Model Based on Observational Registry data from Sweden and Denmark"
author: Erik Bülow,^1,2^ Ute Hahn,^3,4^ Ina Trolle Andersen,^3^ Ola Rolfson,^1,2^  Alma Besic Pedersen,^3,5^ Nils P. Hailer^6^
date: '`r Sys.Date()`'
bibliography: 
#  H:/zotero_lib.bib
  C:\\Users\\erik_\\Documents\\library.bib
csl: clinep.csl
output: 
  bookdown::word_document2:
    reference_docx: article.dotm
    toc: false
    df_print: kable
    number_sections: false
---

@@

- EB: Statistician, MSc, PhD, 0000-0002-9973-456X, erik.bulow@gu.se
- ITA: Statistician, MSc, PhD, 0000-0002-8140-1548, ita@clin.au.dk
- UH: Biostatistician, PhD, associate professor, 000-0001-5059-6293, ute@math.au.dk
- OR: Orthopaedic consultant, PhD, professor, register director, 0000-0001-6534-1242, ola.rolfson@gu.se
- ABP: Orthopaedic consultant, PhD, DMSc, professor, 0000-0002-3288-9401, abp@clin.au.dk
- NPH: Orthopaedic consultant, PhD, professor, 0000-0002-3233-2638, nils.Hailer@surgsci.uu.se

@@

@@

@@

@@

# Conclusion

@@

# Abbreviations

ACS NSQIP	Universal American College of Surgeons National Surgical Quality Improvement Program
ASA		American Society for Anesthesiologists
ATC		Anatomical Therapeutic Chemical
AUC		area under the curve 
BMI		body mass index
CI		Confidence interval
DHR		The Danish Hip Arthroplasty Register
DK		Denmark
FNI		fraction of new information
ICD		International Classification of Diseases
LASSO		least absolute shrinkage and selection operator
LISA		The longitudinal integration database for health insurance and labour market studies
NOMESCO		Nordic Medico-Procedural Committee
NPR		The National Patient Register
PJI		periprosthetic joint infection
R2		Nagelkerke's pseudo-coefficient of determination
ROC		Receiver operating characteristic
SD		Standard deviation
SE		Sweden
SHAR		The Swedish Hip Arthroplasty Register
THA		total hip arthroplasty
WHO		World Health Organization

@@

# Funding

None

# Conflict of interest

@@

```{r comparecoefs}
compare_se_dk_coefs %>% 
  rename(`95 % CI` = CI) %>% 
knitr::kable(
  caption = "Estimated coefficients for the reduced model based on the Swedish derivation cohort (Swedish), as well as re-estimated coefficient values based on the Danish cohort (Danish). Empirical confidence intervals (95 % CI) based on 1,000 resamples from the Swedish cohort of the same sample size (n = 18,854) as used in the Danish cohort. The Danish estimates fall within most of the CIs, indicating no support to reject the null hypothesis of no differneces between the countries."
)
```
